Example Manuscript Template for a Data Analysis Project

### 2024-03-15 products/manuscript/workspace.qmd
Getting this project back on track!
Hopefully getting this project back on track!

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed.

You don't need to have exactly these sections, but the content covering those sections should be addressed.

@@

According to data provided by the World Health organization, seasonal respiratory illnesses are linked to over 650,000 deaths each year worldwide. In the United States the flu is responsible for approximately 42,000 deaths per year and ranks 13th most frequent cause of death.  As such, influenza is the subject of extensive public health surveillance and mediation efforts.

@@

The research questions that I would like to answer are:

1. Can the progression of flu outbreaks be traced geographically or temporally over the course of a year?

2. Is there a specific location from which the annual predominant strain of flu originates?

3. If a pattern can be established, does it repeat on a yearly basis? Does the pattern shift from year to year? Can other factors, such as weather, account for any changes to this pattern?

I am certain that efforts like this already exist. My purpose for this project is to use it as a learning tool and to see if I can replicate those prior efforts.  Data to support this project is readily availble from multiple sources such as GISAID (<https://gisaid.org/>), the World Health Organization Global Influenza Programme (<https://www.who.int/tools/flunet>).

@@

@@

@@

Data was acquired from the World Health Organization Global Influenza Programme. The data includes the excel files containing raw data (VIW_FNT_CM.xlsx) and a file containing the legend for the data (VIW_FLU_METADATA.csv).  This data set is one example of the data available for a project of this sort.  This specific data set includes information on 89 different parameters associated with influenza outbreaks as reported world-wide since about 1995.  The data includes time, region, type of flu, and many opther variables that are important for answereing the questions posed above.

## Data import and cleaning

@@

@@

@@

@@

@@

@@

```
These graphs help whittle down the data a bit.  For example, we can look at the stacked grpah in Figure 4 to see that over the course of the year the major flu variant shifted from influenza A to influenza B.  That being said, this format is not helpful for looking at the incidence over time per country or region.  To look at the sort of data we could do a stacked bar graph of the total flu counts per region/country or perhaps create a graph of the data using a simple scatter plot.

@@

```
Figure 6

## Statistical analysis
_Explain anything related to your statistical analyses._

@@

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@@

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@@

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

@@

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

>>>>

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. You don't need to have exactly these sections, but the content covering those sections should be addressed.

@@

According to data provided by the World Health organization, seasonal respiratory illnesses are linked to over 650,000 deaths each year worldwide. In the United States the flu is responsible for approximately 42,000 deaths per year and ranks 13th most frequent cause of death.

@@

The research questions that i would like to answer are: 1. Can the progression of flu outbreaks be traced geographically and temporally over the course of a year? 2. Is there a specific location from which the annual predominant strain of flu originates? 3. If a pattern can be established, does it repeat on a yearly basis? Does the pattern shift from year to year? Can other factors, such as weather, account for any changes to this pattern?

Note: I am certain that efforts like this already exist. My purpose for this project is to use it as a learning tool and to see if I can relicate those prior efforts.

@@

*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*

@@

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender. We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("assets","antigen-recognition.png"))
```

@@

Data was acquired from the World Health Organization Global Influenza Programme. The data includes the excel files containing raw data (VIW_FNT_CM.xlsx) and a file containing the legend for the data (VIW_FLU_METADATA.csv).

## Data import and cleaning
## Data import and cleaning

@@

# Plot the time axis data using a layered bar graph
plot1 <- ggplot(WHO_data_flu, aes(x = MMWR_WEEKSTARTDATE)) +
  geom_bar(aes(y = INF_ALL, fill = "INF_ALL"), stat = "identity", color="black", position = "stack") + 
  geom_bar(aes(y = INF_A, fill = "INF_A"), stat = "identity", color="black", position = "stack") + 
  geom_bar(aes(y = INF_B, fill = "INF_B"), stat = "identity", color="black", position = "stack") + 
  labs(title = "Influenza Cases Reported to WHO",
       subtitle = "Country = all, 1995-2024 Season",
       x = "Week",
       y = "Number of Positive Specimens",
       fill = "Influenza Type") + 
  theme_minimal()

plot1save = here("assets", "plot1.png")
ggsave(plot1save, plot1, width = 10, height = 6, dpi = 300)

@@

# Plot the region axis data using a layered bar graph
plot1B <- ggplot(WHO_data_flu, aes(x = WHOREGION)) +
  geom_bar(aes(y = INF_ALL, fill = "INF_ALL"), stat = "identity", color="black", position = "stack") + 
  labs(title = "Influenza Cases Reported to WHO",
       subtitle = "Country = all, 1995-2024 Season",
       x = "WHO Region",
       y = "Number of Positive Specimens",
       fill = "Influenza Type") + 
  theme_minimal()

plot1Bsave = here("assets", "plot1B.png")
ggsave(plot1Bsave, plot1B, width = 10, height = 6, dpi = 300)

@@

```
```{r}

WHO_data_flu1 <- WHO_data_flu %>% filter(COUNTRY_AREA_TERRITORY == "China")
WHO_data_flu1 <- WHO_data_flu1 %>% filter(COUNTRY_AREA_TERRITORY == "China")
WHO_data_flu1 <- na.omit(WHO_data_flu1)

library(gt)
WHO_data_flu1 |> 
  gt()

# Plot the data using a layered bar graph
plot2 <- ggplot(WHO_data_flu1, aes(x = MMWR_WEEKSTARTDATE)) +
  geom_bar(aes(y = INF_ALL, fill = "INF_ALL"), stat = "identity", color="black", position = "stack") + 
  geom_bar(aes(y = INF_A, fill = "INF_A"), stat = "identity", color="black", position = "stack") + 
  geom_bar(aes(y = INF_B, fill = "INF_B"), stat = "identity", color="black", position = "stack") + 
  labs(title = "Influenza Cases Reported to WHO",
       subtitle = "Country = China, 2005-2024 Seasons",
       x = "Week",
       y = "Number of Positive Specimens",
       fill = "Influenza Type") + 
  theme_minimal()

plot2save = here("assets", "plot2.png")
ggsave(plot2save, plot2, width = 10, height = 6, dpi = 300)

@@

```{r}
WHO_data_flu2 <- WHO_data_flu1 %>% filter(MMWR_WEEKSTARTDATE > '2023-10-01')
WHO_data_flu2 <- WHO_data_flu2 %>% filter(MMWR_WEEKSTARTDATE > '2023-10-01')
WHO_data_flu2 <- na.omit(WHO_data_flu2)

library(gt)
WHO_data_flu2 |> 
  gt()

# Plot the data using a layered bar graph
plot3 <- ggplot(WHO_data_flu2, aes(x = MMWR_WEEKSTARTDATE)) +
  geom_bar(aes(y = INF_ALL, fill = "INF_ALL"), stat = "identity", color="black", position = "stack") + 
  geom_bar(aes(y = INF_A, fill = "INF_A"), stat = "identity", color="black", position = "stack") + 
  geom_bar(aes(y = INF_B, fill = "INF_B"), stat = "identity", color="black", position = "stack") + 
  labs(title = "Influenza Cases Reported to WHO",
       subtitle = "Country = China, 2022-2024 Seasons",
       x = "Week",
       y = "Number of Positive Specimens",
       fill = "Influenza Type") + 
  theme_minimal()

plot3save = here("assets", "plot3.png")
ggsave(plot3save, plot3, width = 10, height = 6, dpi = 300)

@@

WHO_data_fluB <- WHO_data_flu %>% filter(MMWR_WEEKSTARTDATE > '2023-05-01')
WHO_data_fluB <- WHO_data_fluB %>% filter(MMWR_WEEKSTARTDATE > '2023-05-01')
WHO_data_fluB <- na.omit(WHO_data_fluB)

library(gt)
WHO_data_fluB |> 
  gt()

# Plot the region axis data using a layered bar graph
plot1C <- ggplot(WHO_data_fluB, aes(x = WHOREGION)) +
  geom_bar(aes(y = INF_ALL, fill = "INF_ALL"), stat = "identity", color="black", position = "stack") + 
  labs(title = "Influenza Cases Reported to WHO",
       subtitle = "Country = all, 2022-2024 Season",
       x = "WHO Region",
       y = "Number of Positive Specimens",
       fill = "Influenza Type") + 
  theme_minimal()

plot1Csave = here("assets", "plot1C.png")
ggsave(plot1Csave, plot1C, width = 10, height = 6, dpi = 300)

@@

plot2B <- ggplot(WHO_data_fluB, aes(x=MMWR_WEEKSTARTDATE)) +
  geom_point(aes(y=INF_ALL, color = WHOREGION)) +
  geom_line(aes(y=INF_ALL, color = WHOREGION)) +
    labs(x = "DATE", y = "Total Reported Cases of FLU", title = "Scatterplot Flu Cases Per WHO Region for the 2023-2024 Flu Season") +
  theme_minimal()

plot2Bsave = here("assets", "plot2B.png")
ggsave(plot2Bsave, plot2B, width = 10, height = 6, dpi = 300)

@@

WHO_data_fluB <- WHO_data_flu %>% filter(WHOREGION == "AFR") %>% mutate (INF_AFR = INF_ALL)
WHO_data_fluB <- WHO_data_flu %>% filter(WHOREGION == "AMR") %>% mutate (INF_AMR = INF_ALL)
WHO_data_fluB <- WHO_data_flu %>% filter(WHOREGION == "EUR") %>% mutate (INF_EUR = INF_ALL)
WHO_data_fluB <- WHO_data_flu %>% filter(WHOREGION == "WPR") %>% mutate (INF_WPR = INF_ALL)

plot2A <- ggplot(WHO_data_fluB, aes(x = MMWR_WEEKSTARTDATE)) +
  geom_point(aes(y = INF_AFR), color = "red") +
  geom_line(aes(y = INF_AFR), color = "red") +
  geom_point(aes(y = INF_AMR), color = "blue") +
  geom_line(aes(y = INF_AMR), color = "blue") +
  geom_point(aes(y = INF_EUR), color = "green") +
  geom_line(aes(y = INF_EUR), color = "green") +
  geom_point(aes(y = INF_WPR), color = "magenta") +
  geom_line(aes(y = INF_WPR), color = "magenta") +
  labs(x = "DATE", y = "Total Reported Cases of FLU", title = "Scatterplot Flu Cases Per WHO Region for the 2023-2024 Flu Season") +
  theme_minimal()

plot2Asave = here("assets", "plot2A.png")
ggsave(plot2Asave, plot2A, width = 10, height = 6, dpi = 300)

#| label: fig-schematic
#| fig-cap: "Figure 6: World Flu Burden by Region Over the 2023-2024 Flu Season."
#| echo: FALSE
knitr::include_graphics(here("assets","plot2A.png"))

## Statistical analysis

*Explain anything related to your statistical analyses.*

@@

*Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project.*

@@

*To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*

@@

*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here.*

@@

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*
